---
title: "bhammpsexp"
author: "Umar Hassan"
date: "13 January 2021"
output: html_document
---

##After getting my figures completely wrong for the MPs expenses, I will attempt to rectify this by redoing my analysis for Birmingham-based MPs expenses, finding out which MPs have claimed the most in expenses in Birmingham.

First, let's load the packages I will be using to do my analysis in R and SQL.

```{r}
library(sqldf)
library(tidyverse)
```


```{r}
mpsexp2021 <- read.csv("Individual_claims_for_20_21.csv", stringsAsFactors = FALSE)
mpsexp1920 <- read.csv("Individual_claims_for_19_20.csv", stringsAsFactors = FALSE)
mpsexp1819 <- read.csv("Individual_claims_for_18_19.csv", stringsAsFactors = FALSE)
mpsexp1718 <- read.csv("Individual_claims_for_17_18.csv", stringsAsFactors = FALSE)
mpsexp1617 <- read.csv("Individual_claims_for_19_20.csv", stringsAsFactors = FALSE)

```


#Now, what I need to do next is clean up the colnames using gsub and regex to do this.

```{r}
colnames(mpsexp1617) <- gsub("[., ï]", "",colnames(mpsexp1617))
colnames(mpsexp1718) <- gsub("[., ï]", "",colnames(mpsexp1718))
colnames(mpsexp1819) <- gsub("[., ï]", "",colnames(mpsexp1819))
colnames(mpsexp1920) <- gsub("[., ï]", "",colnames(mpsexp1920))
colnames(mpsexp2021) <- gsub("[., ï]", "",colnames(mpsexp2021))
```

#Great, this will make it easier to analyse the data I will be looking at for MPs expenses in Birmingham.

```{r}
amclbhammps2021 <- sqldf('SELECT MPsName, MPsConstituency, SUM(AmountClaimed)
                         FROM mpsexp2021
                         WHERE MPsConstituency LIKE "%Birmingham%"
                         OR MPsConstituency LIKE "%Sutton Coldfield%"
                         GROUP BY MPsName
                         ORDER BY SUM(AmountClaimed) DESC')
```

#Lovely, I have identified that Khalid Mahmood has topped the list of Birmingham-based MPs who have claimed expenses for the 2020/21 financial year.

Now, let's find out the previous years.

```{r}
amclbhammps1920 <- sqldf('SELECT MPsName, MPsConstituency, SUM(AmountClaimed)
                         FROM mpsexp1920
                         WHERE MPsConstituency LIKE "%Birmingham%"
                         OR MPsConstituency LIKE "%Sutton Coldfield%"
                         GROUP BY MPsName
                         ORDER BY SUM(AmountClaimed) DESC')

```


##The problem I have had is the original figure is not outputting what I want. So, I'll need to check Khalid Mahmood's figure for the same issue.

```{r}
pbmp1920 <- sqldf('SELECT MPsName, MPsConstituency, SUM(AmountClaimed) AS Total
                  FROM mpsexp1920
                  WHERE MPsName LIKE "%Khalid Mahmood%"
                  ORDER BY SUM(AmountClaimed) DESC')

bhammps1920 <- sqldf('SELECT MPsName, MPsConstituency, SUM(AmountClaimed) AS Total
                  FROM mpsexp1920
                  WHERE MPsConstituency LIKE "%Birmingham%"
                  OR MPsConstituency LIKE "%Sutton Coldfield%"
                  GROUP BY MPsName
                  ORDER BY SUM(AmountClaimed) DESC')

bhammps1920 <- sqldf('')

```

